### 2020-06-15 latex/vnhcs.tex
Started writing the preliminaries on mechanical systems and analytical mechanics.

\chapter{Development of Virtual Nonholonomic Constraints}
\section{Mechanical Systems}
\subsection{Lagrangian Formulation of Mechanical Systems}
% TODO: Describe the basics of the lagrangian formulation for mechanical systems
\subsection{Hamiltonian Formulation of Mechanical Systems}
% TODO: Describe how to get the Hamiltonian from the lagrangian, and show that
% you can write the hamiltonian directly in certain conditions of the system
\subsection{Simply Actuated Hamiltonian Systems}
% TODO: Describe the change of coordinates to get into q_u/q_a mode and show
% that the actuator directly affects pa but not pu. Use M for simply actuated
% inertia, D for normal coordinates

>>>>

\chapter{Development of Virtual Nonholonomic Constraints}
\section{Preliminaries on Analytical Mechanics}
In classical mechanics, the position of a point-mass is described using the
three cartesian coordinates (one for each spatial axis) and its motion with the
three corresponding equations of motion (EOM). The position is typically
represented in vector format as a vector in \(\R^3\), while the EOM
are ordinary differential equations derived from Newton's second law 
\(F = m a\).
Generalizing this to a mechanical system with \(N\) point-masses, 
the position of the system can be represented by a vector in \(\R^{3N}\) and the
motion by the \(3N\) EOM. These equations, known as the dynamics of the system, 
are once again derived through Newton's second law \textbf{TODO: SOURCE}.

While the classical approach to mechanics works well for simple systems, solving
for the EOM of a system with more than two particles can be incredibly
challenging - especially if the forces on the system are not known explicitly.
The field of analytical mechanics provides a computational method for finding
these dynamics using only the kinetic and potential energies of the system.

Most mechanical systems are not just a collection of point masses which can 
freely move about three-dimensional space; the points are usually restricted
physically by some equations of constraint (EOC). In a mechanical system with
\(N\) particles with \(m\) EOC, we define the value \(n = 3N - m\) and say the
system has \(n\) degrees of freedom.
Rather than representing the system by one vector in \(\R^{3N}\), it is often
more desirable to represent the position of the system using \(n\)
independent scalar-valued variables called the generalized coordinates.

For example, take two planar spheres linked by an axle as in Figure
\textbf{TODO: draw the points with tikz}. This has 2 point masses at
\((x_1,y_1,z_1)\) and \((x_2,y_2,z_2)\) along with the three EOC 
\((x_2 - x_1)^2 + (y_2 - y_1)^2 - 1 = 0\), \(z_1 = 0\), and \(z_2 = 0\).
The system has 3 degrees of freedom. Instead of using cartesian coordinates, it
is simpler to represent the system with three generalized coordinates:
the \((x,y)\) position of the center of mass and the angle
\(\theta\) between the axle and the horizontal axis.

Generalized coordinates are labelled \(q_i, \, i \in \{1,\ldots,n\}\). 
Each coordinate takes values in \(\Rt{T_i}\), where \(T_i = \infty\) if 
\(q_i\) represents a length, and \(T_i = 2\pi\) if \(q_i\) represents an angle.
It is convention to write \(q = (q_1,\ldots,q_n) \in \mathcal{Q}\) where 
\(\mathcal{Q} = \Rt{T_1} \times \cdot \times \Rt{T_n}\) is called 
the configuration space.

% TODO: Describe the basics of the lagrangian formulation for mechanical systems
% TODO: Describe how to get the Hamiltonian from the lagrangian, and show that
% you can write the hamiltonian directly in certain conditions of the system
\section{Canonical Control Coordinates for Hamiltonian Systems}
% TODO: Describe the change of coordinates to get into q_u/q_a mode and show
% that the actuator directly affects pa but not pu. Use M for simply actuated
% inertia, D for normal coordinates
